---
title: "College Major And Income Data Analysis"
author: "Abdullah Al Asif"
date: "2025-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, message=FALSE}
library(tidyverse)
library(scales)
```


```{r}
recent_grads <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2018/2018-10-16/recent-grads.csv")


majors_processed <- recent_grads %>% 
   arrange(desc(Median)) %>% 
   mutate(Major = str_to_title(Major),
          Major = fct_reorder(Major, Median)) 
```
###Early data Cleaning steps.
###Aggregate by category

```{r  by_major_category}

by_major_category <- majors_processed %>%
   filter(!is.na(Total)) %>% 
  group_by(Major_category) %>%
  summarize(Men = sum(Men),
            Women = sum(Women),
            Total = sum(Total),
            MedianSalary = sum(Median * Sample_size)/ sum(Sample_size)) %>%
  mutate(ShareWomen = Women / Total) %>%
  arrange(desc(ShareWomen))
```

```{r}
glimpse(recent_grads)

```

### Most common majors
what major categories (e.g. Engineering, psychology, business) were most common
```{r}

by_major_category %>% 
   mutate(Major_category = fct_reorder(Major_category, Total)) %>% 
   gather(Gender, Number , Men, Women ) %>% 
   ggplot(aes(Major_category, Number, fill = Gender))+
   geom_col()+
   scale_y_continuous(labels = comma_format())+
   coord_flip()+
   labs(x = "",
        y = "Total # of graduates",
        title =" What are the most common major categories?")


```

###what were The most common *majors*? (since there were 173, we are not going to show them all).


```{r}
majors_processed %>% 
   mutate(Major = fct_reorder(Major, Total)) %>% 
   arrange(desc(Total)) %>% 
   head(20) %>% 
   ggplot(aes(Major, Total, fill = Major_category))+
   geom_col()+
   coord_flip()+
   scale_y_continuous(labels = comma_format())
   labs(x = "",
        y = "Total # of graduates")


```



### what categories of majors make more money than others?




```{r}
majors_processed %>% 
   mutate(Major_category = fct_reorder(Major_category, Median)) %>% 
   ggplot(aes(Major_category, Median, fill = Major_category))+
   geom_boxplot()+
   scale_y_continuous(labels = dollar_format())+
   expand_limits(y = 0)+
   coord_flip()+
   theme(legend.position = "none")
```
```{r}
majors_processed %>% 
   group_by(Major_category) %>% 
  summarize(Median = median(Median))%>% 
   mutate(Major_category = fct_reorder(Major_category, Median)) %>% 
   ggplot(aes(Major_category, Median))+
   geom_col()+
   scale_y_continuous(labels = dollar_format())+
   coord_flip()

```
### What are the highest earning majors?

```{r}


majors_processed %>% 
   filter(Sample_size >= 100) %>% 
   head(20)%>% 
   ggplot(aes(Major, Median, color = Major_category))+
   geom_point()+
   geom_errorbar(aes(ymin = P25th, ymax = P75th))+
   expand_limits(y = 0)+
   coord_flip()+
   scale_y_continuous(labels = dollar_format())+
   labs(title = "What are the highest earning majors?",
          subtitles = "Top 20 Majors with at least 100 Graduates Surveyed. 
        Bars represent 25th to 75th percentitle",x = "" , y = "Median salary of graduates")

```
###what are the lowest Earning Majors?

```{r}
majors_processed %>% 
   filter(Sample_size >= 100) %>% 
   tail(20) %>% 
   ggplot(aes(Major, Median, color = Major_category))+
   geom_point()+
   geom_errorbar(aes(ymin = P25th, ymax = P75th))+
   expand_limits(y = 0)+
   coord_flip()

```

```{r}

majors_processed %>% 
   ggplot(aes(Sample_size, Median))+
   geom_point()+
   geom_text(aes(label = Major), check_overlap = TRUE, vjust = 0, hjust = 0)+
   scale_x_log10()
   
```


#How does gender breakdown relate to typical earnings?#

```{r}
majors_processed %>%
   arrange(desc(Total)) %>% 
   head(20) %>% 
   mutate(Major = fct_reorder(Major, Total)) %>% 
   gather(Gender, Number , Men, Women) %>% 
   select(Major, Gender, Number) %>% 
   ggplot(aes(Major,Number,fill = Gender))+
   geom_col()+
   coord_flip()
   
```



```{r}

library(ggrepel)
by_major_category %>% 
   ggplot(aes(ShareWomen, MedianSalary))+
   geom_point()+
   geom_smooth(method = lm)+
   geom_text_repel(aes(label = Major_category), force = .2)+
   expand_limits(y = 0)
   
```
###interactive Graph

```{r}

library(plotly)


g <- majors_processed %>% 
   mutate(Major_category = fct_lump(Major_category, 4)) %>% 
   ggplot(aes(ShareWomen, Median , color = Major_category, size = Sample_size, label = Major))+
   geom_point()+
   geom_smooth(aes(group = 1),method = lm)+
   scale_x_continuous(labels = percent_format())+
   scale_y_continuous(labels = dollar_format())+
   expand_limits(y = 0)
   

ggplotly(g)
```


```{r}
majors_processed %>% 
   select(Major, Total, ShareWomen, Sample_size, Median) %>% 
   lm(Median ~ ShareWomen, data = ., weights = Sample_size) %>% 
   summary()
```

```{r}
majors_processed %>% 
   filter(Sample_size >=100) %>% 
   mutate(IQR = P75th-P25th) %>% 
   arrange(desc(IQR))

```
```{r}

knitr::knit_exit()

```































































